---
title: "Acute Myocardial Infarction - 30-day mortality"
subtitle: "Data analysis and prognostic model building."
author: "Arhin Claribel Ajua Tawiah, Bragagnolo Sofia, Guccione Anna, Maia Tavora Leticia"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(readxl)
library(tidyverse)
library(ggplot2)
library(scales)
library(epitools)
library(corrplot)
library(gtsummary)
library(knitr)
library(kableExtra)
library(rms)
library(mgcv)
```
# Question 1

> Start with an initial data analysis, comparing patients dead versus patients alive at 30 days. 
Insert also a column with the total population descriptive statistics. Include also a new
variable, the Body Mass Index

## Introduction

**Acute Myocardial Infarction (MI)** (“heart attack”) is caused by the formation 
of a clot in one of the coronary arteries that supply blood to the heart muscle.
Mortality is substantial in the period immediately after the event 
(and during the years after surviving the initial infarction). 

The final aim of this project is to predict the probability of 30-day mortality following an AMI. 

Therefore, the objective of this first part of the analysis is to explore the characteristics of the patients included in the study and compare subjects who were alive at 30 days with those who died within 30 days.

Before fitting regression models, it is important to understand how the variables are distributed, whether there are possible differences between the two groups, and whether the data contain unusual values or missing information. This initial exploration also helps us identify variables that could later become important predictors in the prognostic model.
Another important step is the creation of the Body Mass Index (BMI), which is not directly available in the dataset but can be calculated using height and weight.

## Data loading

```{r load}
data <- read_excel('./AMI_data.xlsx')
```

### Dimensions
The dataset initally includes 3661 patients and 21 variables (20 candidate predictors + 1 outcome variable).
```{r rows}
data.frame(
  'patients'  = nrow(data),
  'factors'   = ncol(data)) %>% 
  kable(col.names = c("# of patients", "# of prognostic factors"), caption = "Dataset dimensions") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("bordered"))
```


### Missing values
Checking for missing values:
```{r missing}
colSums(is.na(data))
```
There is no missing values in the data.

### Dataset description

```{r description}
head(data)
```

```{r str}
str(data)
```

> Although the raw dataset encodes every variable numerically, they fundamentally belong to different statistical types:


| Variable Name | Clinical Description                  | Statistical Type       | Coding / Units in Raw Data                             |   |
|---------------|---------------------------------------|------------------------|--------------------------------------------------------|---|
| day30         | 30-day vital status (Primary Outcome) | Categorical (Binary)   | 0 = Alive, 1 = Dead                                    | 
| sex           | Patient biological sex                | Categorical (Binary)   | 0 = Male, 1 = Female                                   | 
| age           | Age at presentation                   | Numerical (Continuous) | Years                                                  | 
| a65           | Age threshold indicator               | Categorical (Binary)   | 0 = Under 65, 1 = 65 or older                          | 
| killip        | Killip class (Heart failure severity) | Categorical (Ordinal)  | 1 = None, 2 = Rales/S3, 3 = Pulmonary edema, 4 = Shock | 
| sho           | Cardiogenic shock status              | Categorical (Binary)   | 0 = No, 1 = Yes (Killip class 3 or 4)                  | 
| dia           | Diabetes mellitus history             | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| hyp           | Hypotension status                    | Categorical (Binary)   | 0 = No, 1 = Yes (Systolic Blood Pressure < 100 mmHg)   | 
| hrt           | Tachycardia status                    | Categorical (Binary)   | 0 = No, 1 = Yes (Pulse rate > 80 bpm)                  | 
| ant           | Infarction location                   | Categorical (Binary)   | 0 = Other, 1 = Anterior infarct                        | 
| pmi           | Previous myocardial infarction        | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| hei           | Height                                | Numerical (Continuous) | Centimeters (cm)                                       | 
| wei           | Weight                                | Numerical (Continuous) | Kilograms (kg)                                         | 
| htn           | Hypertension history                  | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| smk           | Smoking status                        | Categorical (Nominal)  | 1 = Never, 2 = Ex-smoker, 3 = Current smoker           | 
| lip           | Hypercholesterolaemia                 | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| pan           | Previous angina                       | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| fam           | Family history of MI                  | Categorical (Binary)   | 0 = No, 1 = Yes                                        | 
| ste           | Number of ECG leads with ST elevation | Numerical (Count)      | Number of leads                                        | 
| st4           | Extensive ST elevation threshold      | Categorical (Binary)   | 0 = 4 leads or fewer, 1 = More than 4 leads            | 
| ttr           | Time to pain relief                   | Categorical (Binary)   | 0 = 1 hour or less, 1 = Greater than 1 hour            | 


> Consequently, treating these variables purely as numerical would invalidate regression assumptions. To ensure a clear interpretation these should be converted into factors.

## Data preparation

### BMI variable

We create the BMI variable following the formula:
$BMI=\frac{weight(kg)}{height(m)^2}$

```{r add-bmi}
data$bmi <- data$wei/(data$hei/100)^2
```

### Factors labels

```{r labels}
pretty_labels <- list(
    sex    ~ "Sex",
    age    ~ "Age (years)",
    bmi    ~ "Body Mass Index (kg/m²)",
    killip ~ "Killip class",
    sho    ~ "Shock (Killip 3/4)",
    dia    ~ "Diabetes",
    hyp    ~ "Hypotension (SBP < 100)",
    hrt    ~ "Tachycardia (pulse > 80)",
    ant    ~ "Anterior infarct location",
    pmi    ~ "Previous MI",
    htn    ~ "Hypertension history",
    smk    ~ "Smoking status",
    lip    ~ "Hypercholesterolaemia",
    pan    ~ "Previous angina",
    fam    ~ "Family history of MI",
    ste    ~ "ST elevation (n leads)",
    st4    ~ "ST elevation > 4 leads",
    ttr    ~ "Time to pain relief > 1h"
  )
status_colors <- c("Alive" = "#519e8a", "Dead" = "#E5625E")
```

### Converting to factors

We convert variables into factors for easier interpretation:

```{r day30}
ami_data <- data %>%
  mutate(
    day30 = factor(day30, levels = c(0, 1), labels = c("Alive", "Dead")),
    sex   = factor(sex,   levels = c(0, 1), labels = c("Male", "Female")),
    
    across(
      c(a65, sho, dia, hyp, hrt, ant, pmi, htn, lip, pan, fam, st4, ttr),
      ~ factor(.x, levels = c(0, 1), labels = c("No", "Yes"))
    ),
    
    killip = factor(killip, levels = c(1, 2, 3, 4)),
    smk    = factor(smk, levels = c(1, 2, 3), labels = c("Never", "Ex", "Current"))
  )

head(ami_data)
```

Note that to avoid structural multi-collinearity, a65, sho, hei, wei and st4 are excluded from the final multivariable model.

  * `a65` is a binary age-derived indicator (age ≥ 65) and is therefore excluded to avoid redundancy with continuous variable `age`
  * `hei` and `wei` are used only to compute BMI and are not treated as separate candidate predictors in the predictive model
  * `sho` is derived from Killip (Shock = Killip 3 or 4)
  * `st4` is derived from `ste`

Note that we still included sho and st4 in the exploratory steps.

## Exploring the data
From this initial distribution for a 30-day window, we can see a substantial class imbalance. Surviving patients significantly outnumber deceased patients.
```{r death}
ggplot(ami_data) +
  geom_bar(aes(x = day30, fill = day30), stat = "count") +
  scale_fill_manual(name = "Status", values = status_colors) +
  theme_minimal()
```

### Cumulative incidence
This analysis uses `binom.exact` from the `epitools` package, which calculates the Clopper-Pearson exact binomial confidence interval for the mortality rate.
```{r ci}
n_dead  <- sum(ami_data$day30 == "Dead")
n_total <- nrow(ami_data)
inc <- binom.exact(n_dead, n_total)
data.frame(
  'Deaths'  = n_dead,
  'Total'   = n_total,
  'Perc' = round(inc$proportion * 100, 2),
  'CI lower' = round(inc$lower * 100, 2),
  'CI upper' = round(inc$upper * 100, 2)
) %>% kable(col.names = c("Deaths", "Total cohort (N)", "Incidence (%)", "95% CI lower", "95% CI upper"), caption = "Cumulative 30-Day Mortality Incidence") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("bordered"))
```

The 30-day mortality rate after acute myocardial infarction is relatively low (6.31$\%$), indicating that most patients survived the acute phase. However, this mortality rate remains clinically important because approximately 1 in 16 patients died within 30 days (100/6.31=15.848 $\approx$ 16)

### Outcome balance and events per variable (EPV)
To safely build our prognostic regression model, we must assess the **Events per Variable (EPV)**. For a reliable multivariable logistic regression model that avoids severe overfitting, guidelines generally recommend a minimum baseline of 10 to 15 events (deaths) per estimated parameter.

After excluding redundant and derived variables (`a65`, `sho`, `st4`, `hei`, and `wei`), our final model consists of **16 unique candidate predictors**. 
With 231 total deaths, our dataset yields an EPV of approximately **14.44** per variable. 

```{r epv}
n_predictors <- 16
epv <- n_dead / n_predictors
data.frame(
  'Deaths' = n_dead,
  'Predictors' = n_predictors,
  'EPV' = round(epv, 2),
  'Recommended minimum' = '10-15'
) %>% kable(caption = 'Events Per Variable (EPV) assessment for candidate prognostic predictors') %>%
  kable_styling(full_width = FALSE, bootstrap_options = c('bordered'))

cat(sprintf("\nNote on outcome balance: %.1f%% of the sample are deaths. If the death proportion is below 10–15%%, rare predictors may lead to separation issues in logistic regression.\n", round(n_dead / n_total * 100, 1)))
```

### Descriptive statistics
To compare baseline characteristics between patients who survived and those who died within 30 days, we utilize the Standardized Mean Difference (SMD) rather than traditional p-values.

In large clinical datasets, traditional p-values are highly sensitive to sample size and can flag trivial differences as "highly statistically significant." 
SMD measures the pure effect size and balance independently of sample size.

Interpeting SMD tresholds:
* < 1       negligible difference
* 0.1-0.3   small difference
* 0.3-0.5   moderate difference
* > 0.5     large (clinically substantial) difference

Here we present the table with the characteristics ordered by the absolute value of the SMD.
```{r summary}
tbl_summary(ami_data, by=day30, 
            include = c(age, killip, sho, ant, hrt, ste, st4, smk, hyp, pmi, sex, bmi, pan, ttr, dia, fam, htn, lip),
            label = pretty_labels) %>% add_overall() %>% add_stat_label() %>% bold_labels() %>% add_difference(everything() ~ "smd") %>% 
  modify_spanning_header(all_stat_cols() ~ "**After 30 days**") %>% modify_caption(
    "**Table 1. Baseline characteristics by 30-day vital status**"
  )
```

1. Large Differences (SMD >0.5)
  * Age ($SMD = 60−73= −0.9$): Deceased patients are substantially older (median of 73 years) compared to those who survived (median of 60 years). The negative SMD for age means the "Alive" group has a lower value than the "Dead" group (The Dead group is much older). 
This indicates a substantial difference in age distribution between patients 
who survived and those who died within 30 days.
  * Killip Class (SMD=0.67): Severe heart failure at presentation is a massive indicator of mortality

2. Moderate Differences (SMD 0.3 to 0.5)
  * Deceased patients presented with much higher rates of cardiogenic shock status (sho, 13% vs. 0.9%), anterior infarct locations (ant, 57% vs. 35%), tachycardia (hrt, 51% vs. 30%), and baseline hypotension (hyp, 19% vs. 8.0%).
  * Body Mass Index shows a moderate positive difference. Surviving patients had a higher median BMI (26.9 kg/m²) than those who died (25.3 kg/m²)
  
3. Negligible Differences (SMD <0.1)
  *Hypercholesterolaemia (SMD=0.08): almost perfectly balanced between both groups (40% vs. 36%), indicating that it does not strongly differentiate acute 30-day survival outcomes once the myocardial infarction has already occurred.

### Correlation matrix
We convert our factored variables into numeric values. We include the variables targeted for exclusion (sho, st4, a65) to demonstrate their underlying collinearity. The primary outcome (day30) is excluded as we are only evaluating relationships among the predictors themselves.
For statistical validity, smk was excluded from the matrix layout (as it is a nominal multi-category variable).

The correlation matrix clearly exposes the redundancy of the derived clinical variables, validating the exclusion:
* age and a65 show a high correlation (a65 is a binary split of the continuous age variable)
* killip and sho show a high correlation (confirming that sho is formed by Killip classes 3 and 4)
* ste and st4 show a high correlation (st4 is redundant).

```{r corr_all}
corr_data <- ami_data %>%
  select(age, a65, killip, sho, ant, hrt, ste, st4, hyp, pmi, sex, bmi, pan, ttr, dia, fam, htn, lip) %>%
  mutate(across(everything(), ~ as.numeric(.x)))
corr_matrix <- cor(corr_data, use = "complete.obs")
corrplot(corr_matrix, method = "color", type = "upper", tl.cex = 0.8)
```
## Visual summary
This section presents graphical summaries of key predictors and their relationship with 30-day mortality, helping to identify distributional patterns and group differences before model building.

### Age
From the plots below we can see how older age is strongly associated with 30-day mortality. Patients who died were substantially older, suggesting age is a major prognostic factor after myocardial infarction.

```{r age}
ggplot(ami_data, aes(x = age, fill = day30, colour = day30)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  scale_fill_manual(values  = status_colors) +
  scale_colour_manual(values = status_colors) +
  labs(title = "Age density by 30-day status", x = "Age (years)", y = "Density",
       fill = "30-day status", colour = "30-day status") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

ggplot(ami_data, aes(x = day30, y = age, fill = day30)) +
  geom_violin(alpha = 0.5, trim = TRUE, colour = NA) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, alpha = 0.7) +
  scale_fill_manual(values = status_colors) +
  labs(title = "Age distribution boxplot", x = "Patient status at 30 days", y = "Age (years") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "right", plot.title = element_text(face = "bold"))
```

### Killip

Mortality increases with increasing Killip class, higher Killip class (3,4) reflects more severe heart failure after MI.

We can see that Killip IV has only 19 patients, of which 11 are deaths. 
This suggests potential instability in the estimated odds ratio for Killip IV due to the small sample size.

```{r}
killip_tab <- ami_data %>%
  group_by(killip) %>%
  summarise(
    Alive = sum(day30 == "Alive"),
    Dead = sum(day30 == "Dead"),
    Total = n(),
    Mortality_rate = round(100 * Dead / Total, 1)
  )
killip_tab
```
```{r killip}
ami_data %>%
  count(killip, day30) %>%
  group_by(killip) %>%
  mutate(pct = n / sum(n)) %>%
  filter(day30 == "Dead") %>%
  ggplot(aes(x = killip, y = pct, fill = killip)) +
  geom_col(colour = "white", width = 0.7) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.4, size = 3.8, fontface = "bold") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
  scale_fill_brewer(palette = "Reds") +
  labs(title = "30-day mortality rate by Killip class",
       x = "Killip class", y = "Mortality rate") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none", plot.title = element_text(face = "bold"))
```

### ST elevation
Greater ST segment elevation is clearly associated with a higher probability of death, suggesting that more extensive myocardial involvement worsens an acute prognosis.
```{r ste}
ggplot(ami_data, aes(x = ste)) +
  geom_freqpoly(binwidth = 1, color = "steelblue") +
  labs(title = "ST elevation on ECG (number of leads)") + theme_minimal()

ami_data %>%
  group_by(ste, day30) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ste) %>%
  mutate(pct = n / sum(n)) %>%
  filter(day30 == "Dead") %>%
  ggplot(aes(x = ste, y = pct)) +
  geom_col(fill = "#B71C1C") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mortality by ST elevation in ECG presence",
       x = "ST elevation", y = "Mortality rate") +
  theme_bw(base_size = 11) + theme_minimal()
```
### Sex
Female patients show a higher raw proportion of 30-day mortality compared to male patients.
```{r sex}
ami_data %>%
  group_by(sex) %>%
  summarise(
    Deaths = sum(day30 == "Dead"),
    Total = n(),
    Mortality_Rate = Deaths / Total,
    .groups = "drop"
  ) %>%
  ggplot(aes(x = sex, y = Mortality_Rate, fill = sex)) +
  geom_col(width = 0.5, colour = "white", alpha = 0.8) +
  geom_text(
    aes(label = sprintf("%.1f%%\n(%d / %d)", Mortality_Rate * 100, Deaths, Total)),
    vjust = -0.3, 
    lineheight = 0.9, 
    fontface = "bold", 
    size = 4
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.13)) +
  labs(
    title = "30-day mortality rate by sex",
    x = "sex",
    y = "mortality Rate"
  ) +
  theme_minimal(base_size = 12)
```

### BMI
Patients who died tended to have (surprisingly) lower BMI values.
```{r bmi}
ggplot(ami_data, aes(x = bmi, fill = day30, colour = day30)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  scale_fill_manual(values  = status_colors) +
  scale_colour_manual(values = status_colors) +
  labs(title = "BMI distribution", x = "BMI (kg/m²)", y = "Density",
       fill = "30-day status", colour = "30-day status") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold")) + theme_minimal()

ggplot(ami_data, aes(x = day30, y = bmi, fill = day30)) +
  geom_violin(alpha = 0.5, trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, alpha = 0.7) +
  scale_fill_manual(values = status_colors) +
  labs(title = "BMI by 30-day status", x = NULL, y = "BMI") +
  theme_minimal(base_size = 11) +
  theme(legend.position = "none") + theme_minimal()
```

### Hypotension
Patients presenting with hypotension had substantially higher mortality rates.
```{r}
ami_data %>%
  count(hyp, day30) %>%
  group_by(hyp) %>%
  mutate(rate = n / sum(n)) %>%
  filter(day30 == "Dead") %>%
  ggplot(aes(x = hyp, y = rate, fill = hyp)) +
  geom_col() +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(
    title = "30-day mortality by hypotension status",
    x = "Hypotension",
    y = "Mortality rate"
  )
```

# Question 2

> Perform univariable regression analyses (crude association measures) of all candidate
predictors for your prognostic model. For the continuous variables is the linear effect
reasonable?

## Crude odds and univariable analysis
The table below displays the crude associations between each predictor and 30-day mortality. 

* **Strongest Predictors:** **Killip Class IV** exhibits the highest point estimate, increasing the odds of death nearly 29-fold (**OR = 28.91**, 95% CI: 11.5–75.6) compared to Class I. This is mirrored by the **Shock** variable (**OR = 15.8**). **Hypotension** nearly triples the risk (**OR = 2.79**), and **Tachycardia** more than doubles it (**OR = 2.39**).

* **Anterior infarct location** more than doubles mortality odds (**OR = 2.43**). For **ST elevation**, every additional lead increases the odds of death by 20% (**OR = 1.20**), while having > 4 leads elevated (**st4**) doubles the risk (**OR = 2.01**).
* **Age** increases the odds of death by 10% per year (**OR = 1.10**). **Female sex** (**OR = 1.98**), **Prior MI** (**OR = 2.24**), and **Current smoking** (**OR = 2.23**) all roughly double the crude odds of mortality. **BMI** shows an inverse relationship (**OR = 0.92**), indicating that a lower body mass is associated with higher acute mortality.

```{r crude-or}
tbl_uvregression(
  ami_data,
  label = pretty_labels,
  method   = glm,
  y        = day30,
  method.args = list(family = binomial),
  exponentiate = TRUE,
  include  = c(age, killip, sho, ant, hrt, ste, st4, smk, hyp, pmi, sex, bmi, pan, ttr, dia, fam, htn, lip)
) %>% bold_labels()

tbl_uvregression
```

## Evaluating the linearity assumption for continuous variables
To determine if a linear effect is reasonable for our continuous predictors, we examine how the risk behaves across their ranges on the log-odds scale:
$$
log(\frac{p}{1-p}) = \beta_0+\beta_1X 
$$
We formally test this assumption by comparing a standard linear model against a non-linear model utilizing restricted cubic splines with 4 knots (rcs(..., 4)). A Likelihood Ratio Test (LRT) is then applied to assess whether the spline model offers a statistically significant improvement in fit.

Age, ST Elevation, and BMI are the only true continuous variables in our dataset. 
The other variables are binary or categorical and do not require this linearity assessment.
```{r lin}
# Lspline / restricted cubic splines approach via rms / glm
dd <- datadist(ami_data)
options(datadist = "dd")

check_linearity <- function(var, label) {
  f_lin  <- glm(day30 ~ get(var),                          data = ami_data, family = binomial)
  f_spl  <- glm(day30 ~ rcs(get(var), 4),                  data = ami_data, family = binomial)
  p_val  <- anova(f_lin, f_spl, test = "LRT")$`Pr(>Chi)`[2]
  cat(sprintf("%-6s — LRT linearity p = %.3f  %s\n",
              label, p_val,
              ifelse(p_val < 0.05, "non-linear", "linear OK")))
}

check_linearity("age", "Age")
check_linearity("ste", "ST elevation")
check_linearity("bmi", "BMI")
```

Age (p = 0.020): The LRT rejects strict linearity.
ST Elevation (p = 0.094): The non-significant p-value confirms the linearity assumption.
BMI (p = 0.001): Linearity is strongly rejected.

### GAM Plots

To visually inspect the log-linear relationship assumed by logistic regression, we fit a Generalized Additive Model (GAM) using a penalized regression spline (s()) for each continuous variable. 

```{r vis}
plot_linearity <- function(var, xlab) {
  fit <- gam(as.numeric(day30) - 1 ~ s(get(var)), data = ami_data, family = binomial)
  plot(fit, se = TRUE, rug = TRUE,
       xlab = xlab, ylab = "Log-odds (30-day death)",
       main = paste("GAM smooth —", xlab))
  abline(h = 0, lty = 2)
}

par(mfrow = c(1, 3))
plot_linearity("age", "Age (years)")
plot_linearity("ste", "ST elevation (leads)")
plot_linearity("bmi", "BMI (kg/m²)")
par(mfrow = c(1, 1))
```

The lines represent the estimated smooth effect on the log-odds of 30-day mortality, the dashed lines indicate the 95% confidence intervals, and the the bottom displays the density of individual patient observations.

* **Age** The relationship is clearly curvilinear. Below the age of 50, the risk curve is relatively flat or shallow. However, after age 50, the log-odds of death accelerate sharply upward. Forcing a single straight line through this would underestimate the risk for elderly patients and overestimate the risk for middle-aged patients. This visually justifies using a non-linear term (like a spline).

* **ST Elevation** The smooth function closely approximates a steady, upward straight line across the vast majority of the data distribution (from 0 to 8 leads). While the confidence intervals widen slightly at the extreme upper limit due to sparse data (fewer patients with more than 9 leads elevated), a simple linear term can capture this relationship.

* **BMI** The risk follows a distinct **U-shaped or J-shaped curve**. The log-odds of death drop as BMI increases, reaching the lowest risk around 35 $\text{kg/m}^2$. Past 40 $\text{kg/m}^2$, the risk trajectory begins to curve back upward.

### Final Modeling Decision
Based on the combination of the formal Likelihood Ratio Tests and these GAM diagnostic plots, our final prognostic model will treat **ST Elevation** as a standard linear predictor, while **Age** and **BMI** will be modeled using non-linear transformations.
