---
title: "Acute Myocardial Infarction - 30-day mortality"
subtitle: "Data analysis and prognostic model building."
author: "Arhin Claribel Ajua Tawiah, Bragagnolo Sofia, Guccione Anna, Maia Tavora Leticia"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    highlight: tango
    number_sections: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
library(readxl)
library(epiR)
library(epitools)
library(Epi)
library(popEpi)
library(ggplot2)
library(scales)
library(lubridate)
library(AER)
library(MASS)
library(survival)
library(tidyverse)
library(pmsampsize)
library(psych)
library(corrplot)
library(gtsummary)
library(broom.helpers)
library(smd)
library(knitr)
library(kableExtra)
library(corrplot)
library(rms)
```


**Acute Myocardial Infarction (MI)** (“heart attack”) is caused by the formation of a clot in one of the coronary arteries that supply blood to the heart muscle.
Mortality is substantial in the period immediately after the event (and during the years after surviving the initial infarction). The risk of **30-day mortality** strongly depends on various prognostic factors.

The data we use here are a sub-sample coming from a trial, a four-arm randomized study that
compared the efficacy of different intravenous thrombolytic regimens for acute myocardial infarction.

Aim of this project is to predict the probability of 30-days mortality.

# Question 1: Initial Data Analysis

> Start with an initial data analysis, comparing patients dead versus patients alive at 30 days. 
> Insert also a column with the total population descriptive statistics. 
> Include also a new variable, the Body Mass Index.

## Data loading

```{r load}
ami_data <- read_excel('./AMI_data.xlsx')
```

### Dimensions
```{r rows}
data.frame(
  'patients'  = nrow(ami_data),
  'factors'   = ncol(ami_data)) %>% kable(col.names = c("Num of patients", "Num of prognostic factors"), caption = "Dataset dimensions") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("bordered"))
```

All variables are numerical

```{r description}
head(ami_data)
```

### Factors labels
The factors in the observation include:
  
  * personal factors such as age, gender, height and weight
  * conditions like hypertension, diabetes, smoking
  * family history and the history of previous cardiovascular events (previous MI, angina, stroke, bypass surgery)
  * relevant characteristics such as the location of the infarction and the extent of ECG abnormalities 
  
  * the acute state of the patient as reflected by the Killip score

```{r labels}
pretty_labels <- list(
    sex    ~ "Sex",
    age    ~ "Age (years)",
    bmi    ~ "Body Mass Index (kg/m²)",
    killip ~ "Killip class",
    sho    ~ "Shock (Killip 3/4)",
    dia    ~ "Diabetes",
    hyp    ~ "Hypotension (SBP < 100)",
    hrt    ~ "Tachycardia (pulse > 80)",
    ant    ~ "Anterior infarct location",
    pmi    ~ "Previous MI",
    htn    ~ "Hypertension history",
    smk    ~ "Smoking status",
    lip    ~ "Hypercholesterolaemia",
    pan    ~ "Previous angina",
    fam    ~ "Family history of MI",
    ste    ~ "ST elevation (n leads)",
    st4    ~ "ST elevation > 4 leads",
    ttr    ~ "Time to pain relief > 1h"
  )
status_colors <- c("Alive" = "#519e8a", "Dead" = "#E5625E")
```

### Missing values
Checking for missing values:
```{r missing}
colSums(is.na(ami_data))
```
There is no missing values in the data.

## Data preparation

### BMI factor

We create the BMI column following the formula.

```{r add-bmi}
ami_data$bmi <- ami_data$wei/(ami_data$hei/100)^2
```

### Converting to factors

We convert the `day30` column into a factor for easier interpretation. We do the same with all the other binary variables. And for ordinal variables too.

```{r day30}

ami_data <- ami_data %>%
  mutate(
    day30 = factor(day30, levels=c(0,1), labels=c("Alive","Dead")),
    sex = factor(sex, levels=c(0, 1), labels=c("Male", "Female")),
    a65 = factor(a65, levels=c(0, 1), labels=c("No", "Yes")),
    sho = factor(sho, levels=c(0, 1), labels=c("No","Yes")),
    dia = factor(dia,   levels=c(0,1), labels=c("No","Yes")),
    hyp = factor(hyp,   levels=c(0,1), labels=c("No","Yes")),
    hrt = factor(hrt,   levels=c(0,1), labels=c("No","Yes")),
    ant = factor(ant,   levels=c(0,1), labels=c("No","Yes")),
    pmi = factor(pmi,   levels=c(0,1), labels=c("No","Yes")),
    htn = factor(htn,   levels=c(0,1), labels=c("No","Yes")),
    lip = factor(lip,   levels=c(0,1), labels=c("No","Yes")),
    pan = factor(pan,   levels=c(0,1), labels=c("No","Yes")),
    fam = factor(fam,   levels=c(0,1), labels=c("No","Yes")),
    st4 = factor(st4,   levels=c(0,1), labels=c("No","Yes")),
    ttr = factor(ttr,   levels=c(0,1), labels=c("No","Yes")),
    killip = factor(killip, levels=c(1,2,3,4)),
    smk = factor(ami_data$smk, levels=c(1,2,3), labels=c("Never","Ex","Current"), ordered=TRUE))

head(ami_data)
```

## Exploring the data
```{r death}
ggplot(ami_data) +
  geom_bar(aes(x = day30, fill = day30), stat = "count") +
  scale_fill_manual(name = "Status", values = status_colors) +
  theme_minimal()
```


### Cumulative incidence
```{r ci}
n_dead  <- sum(ami_data$day30 == "Dead")
n_total <- nrow(ami_data)
inc <- binom.exact(n_dead, n_total)
data.frame(
  'Deaths'  = n_dead,
  'Total'   = n_total,
  'Perc' = round(inc$proportion * 100, 2),
  'CI lower' = round(inc$lower * 100, 2),
  'CI upper' = round(inc$upper * 100, 2)
) %>% kable(col.names = c("30-day status", "N", "%", "95% CI lower", "95% CI upper"), caption = "Cumulative incidence at 30 days") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("bordered"))
```

### Descriptive statistics

We add the SMD to confront the variables.
* SMD positive: the dead group has higher values
* SMD negative: the alive group has higher values
* SMD = 0: the 2 groups are identical for that variable
Values:
* < 0.1: negligible
* 0.1 - 0-3: small difference
* 0.3 - 0.5: moderate difference
* > 0.5: large difference

Here we present the table with the characteristics ordered by the absolute value of the SMD.
```{r summary}
tbl_summary(ami_data, by=day30, 
            include = c(age, killip, sho, ant, hrt, ste, st4, smk, hyp, pmi, sex, bmi, pan, ttr, dia, fam, htn, lip),
            label = pretty_labels) %>% add_overall() %>% add_stat_label() %>% bold_labels() %>% add_difference(everything() ~ "smd") %>% modify_spanning_header(all_stat_cols() ~ "**After 30 days**") %>% modify_caption(
    "**Table 1. Baseline characteristics by 30-day vital status**"
  )
```

### Correlation matrix

```{r corr}
small_ami_data <- ami_data %>%
  select(day30, age, killip, sho, ant, hrt, ste, st4, smk, sex, bmi) %>% mutate(across(everything(), ~ as.numeric(.x) - 1))
corr <- cor(small_ami_data, use = "complete.obs")
corrplot(corr, method = "color", type = "upper")
```
## Visual summary
Now we present some visualization for key predictors.



### Age

```{r age}
# Histogram for age
ggplot(ami_data, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(title = "Age distribution") + theme_minimal()

ggplot(ami_data, aes(x = age, fill = day30, colour = day30)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  scale_fill_manual(values  = status_colors) +
  scale_colour_manual(values = status_colors) +
  labs(title = "Age distribution by 30-day status", x = "Age (years)", y = "Density",
       fill = "30-day status", colour = "30-day status") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold")) + theme_minimal()

ggplot(ami_data, aes(x = day30, y = age, fill = day30)) +
  geom_violin(alpha = 0.5, trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, alpha = 0.7) +
  scale_fill_manual(values = status_colors) +
  labs(title = "Age by 30-day status", x = NULL, y = "Age") +
  theme_bw(base_size = 11) +
  theme(legend.position = "none") + theme_minimal()
```

### Killip

```{r killip}
ami_data %>%
  count(killip, day30) %>%
  group_by(killip) %>%
  mutate(pct = n / sum(n)) %>%
  filter(day30 == "Dead") %>%
  ggplot(aes(x = killip, y = pct, fill = killip)) +
  geom_col(colour = "white", width = 0.7) +
  geom_text(aes(label = scales::percent(pct, accuracy = 0.1)),
            vjust = -0.4, size = 3.8, fontface = "bold") +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.65)) +
  scale_fill_brewer(palette = "Reds") +
  labs(title = "30-day mortality rate by Killip class",
       x = "Killip class", y = "Mortality rate") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none", plot.title = element_text(face = "bold")) + theme_minimal()
```

### ST elevation

```{r ste}

ggplot(ami_data, aes(x = ste)) +
  geom_freqpoly(binwidth = 1, color = "steelblue") +
  labs(title = "ST elevation on ECG (number of leads)") + theme_minimal()

ami_data %>%
  group_by(ste, day30) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(ste) %>%
  mutate(pct = n / sum(n)) %>%
  filter(day30 == "Dead") %>%
  ggplot(aes(x = ste, y = pct)) +
  geom_col(fill = "#B71C1C") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mortality by ST elevation in ECG presence",
       x = "ST elevation", y = "Mortality rate") +
  theme_bw(base_size = 11) + theme_minimal()
```

### Sex
```{r sex}
# Bar chart: sex + outcome
ggplot(ami_data, aes(x = factor(sex), fill = day30)) +
  scale_fill_manual(name = "Status", values = status_colors) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Sex distribution and status after 30 days", x = "Sesso (0=M, 1=F)") + theme_minimal()
```

### BMI

```{r bmi}
ggplot(ami_data, aes(x = bmi, fill = day30, colour = day30)) +
  geom_density(alpha = 0.4, linewidth = 0.8) +
  scale_fill_manual(values  = status_colors) +
  scale_colour_manual(values = status_colors) +
  labs(title = "BMI distribution", x = "BMI (kg/m²)", y = "Density",
       fill = "30-day status", colour = "30-day status") +
  theme_bw(base_size = 12) +
  theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

# Violinplot: BMI + outcome
ggplot(ami_data, aes(x = day30, y = bmi, fill = day30)) +
  geom_violin(alpha = 0.5, trim = TRUE) +
  geom_boxplot(width = 0.15, outlier.size = 0.5, alpha = 0.7) +
  scale_fill_manual(values = status_colors) +
  labs(title = "BMI by 30-day status", x = NULL, y = "BMI") +
  theme_bw(base_size = 11) +
  theme(legend.position = "none") + theme_minimal()
```


# Question 2

> Perform univariable regression analyses (crude association measures) of all candidate
predictors for your prognostic model. 
> For the continuous variables is the linear effect reasonable?

## Crude odds
```{r crude-or}
tbl_uvregression(
  ami_data,
  label = pretty_labels,
  method   = glm,
  y        = day30,
  method.args = list(family = binomial),
  exponentiate = TRUE,
  include  = c(age, killip, sho, ant, hrt, ste, st4, smk, hyp, pmi, sex, bmi, pan, ttr, dia, fam, htn, lip)
) %>% bold_labels()
```

## Linearity check for continuous predictors


```{r lin}
# Lspline / restricted cubic splines approach via rms
dd <- datadist(ami_data); options(datadist = "dd")

check_linearity <- function(var, label) {
  f_lin  <- glm(day30 ~ get(var),                          data = ami_data, family = binomial)
  f_spl  <- glm(day30 ~ rcs(get(var), 4),                  data = ami_data, family = binomial)
  p_val  <- anova(f_lin, f_spl, test = "LRT")$`Pr(>Chi)`[2]
  cat(sprintf("%-6s — LRT linearity p = %.3f  %s\n",
              label, p_val,
              ifelse(p_val < 0.05, "⚠ non-linear", "linear OK")))
}

check_linearity("age", "Age")
check_linearity("ste", "ST elevation")
check_linearity("bmi", "BMI")
check_linearity("hei", "Height")
check_linearity("wei", "Weight")
```

### Visual check

```{r vis}
library(mgcv)

plot_linearity <- function(var, xlab) {
  fit <- gam(as.numeric(day30) - 1 ~ s(get(var)), data = ami_data, family = binomial)
  plot(fit, se = TRUE, rug = TRUE,
       xlab = xlab, ylab = "Log-odds (30-day death)",
       main = paste("GAM smooth —", xlab))
  abline(h = 0, lty = 2)
}

par(mfrow = c(1, 3))
plot_linearity("age", "Age (years)")
plot_linearity("ste", "ST elevation (leads)")
plot_linearity("bmi", "BMI (kg/m²)")
plot_linearity("hei", "Height (m)")
plot_linearity("wei", "Weight (kg)")
par(mfrow = c(1, 1))
```
